\pdftitle

### 2017-05-14 maxNt.tex
add discrepancy section

%% Author: PGL  Porta Mana
%% Created: 2015-11-04T09:20:08+0100
%% Last-Updated: 2017-05-15T00:15:08+0200
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Report-no: ***
\documentclass{article}

@@

It is at this point that the \me\ method enters the scene. It selects one
distribution, purported to be \enquote{maximally noncommittal}, among those
that have the desired expectations. But we are in a quandary:
formulae~\eqref{eq:expe_products} allow us to apply the method to find the
probability of the population $\p(\yXf=\yRf \| \yH)$, or of the sample
$\p(\yxs=\yrs \| \yH)$; the two applications, however, are
\emph{inequivalent}. They lead to numerically different
$\p(\yxs=\yrs \| \yH)$; see \fig~\ref{fig:diff_maxent_pop_sample}. Which
should we choose?

\begin{figure}[!b]
\centering
\includegraphics[width=0.95\columnwidth]{different_maxent_pop_sample.pdf}%
\caption{Example of discrepancy in the probability $\p(\yxs = \yrs \| \yH)$
  constructed by \me\ at the population level followed by sample
  marginalization (blue triangles), and at the sample level (red circles).
  The constraints are the same in both applications:
  $\expe{\yxs \|\yH}=0.45$, $\expe{\yxxs \|\yH}=0.35$, and $N=5\,000$,
  $n=200$.}
\label{fig:diff_maxent_pop_sample}
\end{figure}%maxent_pop_or_sample.nb
\iffalse
\begin{align}
  \expe{x_{i_1} \dotsm x_{i_m} \| \yH}
&=
\expe{\sav{\underbrace{\yx \dotsm \yx}_{\text{$m$ factors}}} \| \yH}
=
\expe{\av{\underbrace{\yX \dotsm \yX}_{\text{$m$ factors}}} \| \yH},
\\
\intertext{and has an explicit expression in terms of $Q$:}
  \begin{split}
\expe{x_{i_1} \dotsm x_{i_m} \| \yH}
&=
\binom{N}{m}^{-1}
%\frac{(N-m)!}{N!}
\sum_{N A=0}^N %m!\, 
\binom{N A}{m}\, Q(A)
\\
&\equiv
%\frac{(N-m)!}{N!}
\sum_{N A=0}^N %m!\, 
\binom{N-m}{N A-m}
\binom{N}{N A}^{-1} Q(A).
\end{split}
\end{align}
\end{subequations}
\fi
***

>>>>

%% Author: PGL  Porta Mana
%% Created: 2015-11-04T09:20:08+0100
%% Last-Updated: 2017-05-15T01:20:54+0200
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Report-no: ***
\documentclass{article}

@@

It is at this point that the \me\ method enters the scene. It selects one
distribution, purported to be \enquote{maximally noncommittal}, among those
that have the desired expectations. But we are in a quandary:
formulae~\eqref{eq:expe_products} allow us to apply the method to find the
probability of the population $\p(\yXf=\yRf \| \yH)$, or of the sample
$\p(\yxs=\yrs \| \yH)$; the two applications, however, are
\emph{inequivalent}. They lead to numerically different
$\p(\yxs=\yrs \| \yH)$; see \fig~\ref{fig:diff_maxent_pop_sample} for an
example. Suppose our constraints are $\expe{\yxs \|\yH}$ and
$\expe{\yxxs \|\yH}$. Application of \me\ at the population level leads to
the distribution
\begin{equation}
  \label{eq:full_pop_av_example_ME}
  \p(\yXf = \yRf \| \yH)  \propto
\binom{N}{N \yRf}\exp\biggl[
\yL_2\, \frac{\yRf\,(N \yRf-1)}{N-1} + \yL_1\, \yRf\biggr]
\end{equation}
with such parameters $\yL_1$, $\yL_2$ as to satisfy the constraints.
Marginalization by \eqn~\eqref{eq:marginal} leads to 
\begin{equation}
  \label{eq:full_pop_av_example_ME_marg}
  \p(\yxs = \yrs \| \yH)  \propto
  \sum_{N\yRf=0}^N
  \binom{n}{n\yrs} \binom{N-n}{N\yRf-n\yrs} \exp\biggl[
\yL_2\, \frac{\yRf\,(N \yRf-1)}{N-1} + \yL_1\, \yRf\biggr].
\end{equation}
Application of \me\ at the sample level leads to the distribution
\begin{equation}
  \label{eq:sample_av_example_ME}
  \p(\yxs = \yrs \| \yH)  \propto
  \binom{n}{n\yrs}
  \exp\biggl[\yl_2\, \frac{\yrs\,(n \yrs-1)}{n-1} + \yl_1\, \yrs\biggr]
\end{equation}
with such parameters $\yl_1$, $\yl_2$ as to satisfy the constraints.
Expressions \eqref{eq:full_pop_av_example_ME_marg} and
\eqref{eq:sample_av_example_ME} cannot be equal when $n>6$ because, even if
we could choose their four parameters and two normalization constants
freely (and we can't), for their equality we should still satisfy $n+1$ equations,
\ie\ more than six.

Which should we choose?

\begin{figure}[!t]
\centering
\includegraphics[width=0.95\columnwidth]{different_maxent_pop_sample_100.pdf}%
\caption{Example of discrepancy in $\p(\yxs = \yrs \| \yH)$ constructed by
  \me\ at the population level followed by sample marginalization (blue
  triangles), and at the sample level (red circles). The constraints are
  the same in either case: $\expe{\yxs \|\yH}=0.45$,
  $\expe{\yxxs \|\yH}=0.35$, and $N=5\,000$, $n=100$.}
\label{fig:diff_maxent_pop_sample}
\end{figure}%maxent_pop_or_sample.nb
\iffalse
\begin{align}
  \expe{x_{i_1} \dotsm x_{i_m} \| \yH}
&=
\expe{\sav{\underbrace{\yx \dotsm \yx}_{\text{$m$ factors}}} \| \yH}
=
\expe{\av{\underbrace{\yX \dotsm \yX}_{\text{$m$ factors}}} \| \yH},
\\
\intertext{and has an explicit expression in terms of $Q$:}
  \begin{split}
\expe{x_{i_1} \dotsm x_{i_m} \| \yH}
&=
\binom{N}{m}^{-1}
%\frac{(N-m)!}{N!}
\sum_{N A=0}^N %m!\, 
\binom{N A}{m}\, Q(A)
\\
&\equiv
%\frac{(N-m)!}{N!}
\sum_{N A=0}^N %m!\, 
\binom{N-m}{N A-m}
\binom{N}{N A}^{-1} Q(A).
\end{split}
\end{align}
\end{subequations}
\fi
***
